---
output:
  html_document: default
  pdf_document: default
---
---
title: 'ggplot2 application (car:: Salaries)'
author: "cafepeng"
date: "2018年3月18日"
output: html_document
helpful link: [https://support.zendesk.com/hc/en-us/articles/203691016-Formatting-text-with-Markdown#topic_xqx_mvc_43__row_tf4_bmn_1n]
-------------------------
**Dataset:**

> The 2008-09 nine-month academic salary for Assistant Professors, Associate Professors and Professors in a college in the U.S. 
The data were collected as part of the on-going effort of the college's administration to monitor salary differences between male and female faculty members.

* Variables:
    * rank: AssocProf, AsstProf, Prof
    * discipline:a factor with levels A (`theoretical')or B ('applied' departments)
    * yrs.since.phd: years since PhD
    * yrs.service: years of service
    * sex: a factor with levels Female Male
    * salary: nine-month salary, in dollars


1. pull the dataset "Salaries" from 'car' package, and check the structure and so on.
```{r}
library(car)
data01<-data.frame(Salaries)
str(data01)
dim(data01)
summary(data01)

#especially check the variables I am interested in
summary(data01$rank)
summary(data01$salary)
summary(data01$sex)

```


* **Q1. If income increases as the year of service increases? **
```{r}
#Load the packages needed for data visualization (note: eval=FALSE prevents code from being evaluated)
library(ggplot2)
#note: alpha refers to the degree of transparancy so that the points would not block other points
pic1<-ggplot(data01)+geom_point(mapping=aes(x=yrs.service,y=salary),alpha=1/2,color="blue",size=5)
pic1

```

**Q1 Finding: as years of service increases, the salaries increase as well**


* **Q2.is there a difference in salaries between males and females overall?**
```{r}
# add another aesthetic measure "sex"
library(ggplot2)
pic2<-ggplot(data01)+geom_point(mapping=aes(x=yrs.service,y=salary,color=sex),alpha=1/2,size=5)
pic2
```

**Q2 Findings: it seems that the salary range is relatively limited to females**

* **Q3_1. following Q2, is it possible that the difference btw males and females gets larger in higher income range(e.g. higher than Q3)?**
```{r}
library(ggplot2)
pic3_1<-ggplot(data01)+geom_point(mapping=aes(x=yrs.service,y=salary>134185,color=sex),alpha=1/2,size=5)
pic3_1
```

**Q3-1 Finding: due to the relative small sample of female, this picture is relatively meaningless**


* **Q3_2.Use "facet"" setting to draw two diagrams that represents females and males seperately**
```{r}
library(ggplot2)
pic3_2<-ggplot(data01)+geom_point(mapping=aes(x=yrs.service,y=salary),color="blue",alpha=1/2,size=5)+facet_wrap(~sex)
pic3_2
```

**Q3_2 Finding: females' salary range tends to be relatively limited compared to males**

* **4_1.Does the salary differs in terms of ranks?**
```{r}
library(ggplot2)
pic4_1<-ggplot(data01,mapping=aes(x=yrs.service,y=salary))+
  geom_point(
      mapping=aes(color=rank),alpha=1/2,size=5)+
  geom_smooth(
      mapping=aes(color=rank))
pic4_1
```

**Q4_1 Finding: higher the rank, higher the variance in salaries**

*Note: we can use the command 'ggsave' to save the diagram. `ggsave(filename="pic4_1.pdf",plot=pic4_1)`*


* **Q4_2: Take a closer look at the highest rank "professor" by plotting its line only**

```{r}
library(ggplot2)
require(dplyr) # filter is part of dplyr package
pic4_2<-ggplot(data01,mapping=aes(x=yrs.service,y=salary))+geom_point(mapping=aes(color=rank),alpha=1/2,size=5)+geom_smooth(data=filter(data01,rank=="Prof"), se=FALSE)
pic4_2
```

**Q4_2 Finding: higher the rank, highe the range in salaries**
